% WACV 2026 Paper Template
% based on the ICCV 2025 template (https://media.eventhosts.cc/Conferences/ICCV2025/ICCV2025-Author-Kit-Feb.zip) with
% WACV-specific details (e.g., 2 tracks) from the WACV 2025 template (https://www.dropbox.com/scl/fi/su44zgdhrzik26p2xu37k/WACV-2025-Author-Kit-Template.zip?rlkey=5qcfimjhxnmx3wlyk7yhk8wg7&dl=0)

\documentclass[10pt,twocolumn,letterpaper]{article}

%%%%%%%%% PAPER TYPE  - PLEASE UPDATE FOR FINAL VERSION
\usepackage[review,algorithms]{wacv}      % To produce the REVIEW version for the algorithms track
%\usepackage[review,applications]{wacv}      % To produce the REVIEW version for the applications track
%\usepackage{wacv}              % To produce the CAMERA-READY version
%\usepackage[pagenumbers]{wacv} % To force page numbers, e.g. for an arXiv version

% Import additional packages in the preamble file, before hyperref
\input{preamble}

% It is strongly recommended to use hyperref, especially for the review version.
% hyperref with option pagebackref eases the reviewers' job.
% Please disable hyperref *only* if you encounter grave issues, 
% e.g. with the file validation for the camera-ready version.
%
% If you comment hyperref and then uncomment it, you should delete *.aux before re-running LaTeX.
% (Or just hit 'q' on the first LaTeX run, let it finish, and you should be clear).
\definecolor{wacvblue}{rgb}{0.21,0.49,0.74}
\usepackage[breaklinks,colorlinks,allcolors=wacvblue]{hyperref}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\wacvPaperID{*****} % *** Enter the WACV Paper ID here
\def\confName{WACV}
\def\confYear{2026}


%%%%%%%%% TITLE - PLEASE UPDATE
\title{Depth-Guided and Tree-Focused Learning for Tree Classification under Complex Backgrounds}

%%%%%%%%% AUTHORS - PLEASE UPDATE
\author{line 1: 1st Given Name Surname\\
line 2: dept. name of organization\\
(of Affiliation)\\
line 3: name of organization\\
(of Affiliation)\\
line 4: City, Country\\
line 5: email address or ORCID}

\begin{document}
\maketitle
\input{sec/0_abstract}
\input{sec/1_introduction}
\input{sec/2_related_work}
\input{sec/3_method}
\input{sec/4_datasets}
\input{sec/5_results}
\input{sec/6_discussion}
\input{sec/7_conclusion}
\input{sec/8_futurework}
% References (thebibliography; switch to BibTeX later if needed)
\begingroup
\small
\begin{thebibliography}{99}
\bibitem{ref1} A. Dosovitskiy et al., ``An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale,'' in Proc. Int. Conf. Learn. Representations (ICLR), 2021.
\bibitem{ref2} J. Peters, S. Wild, and L. Schmidt, ``Attention-Based Multiple Instance Learning for Visual Recognition,'' in Proc. Int. Conf. Mach. Learn. (ICML), 2019, pp. 7745–7755.
\bibitem{ref3} R. Ranftl, K. Lasinger, D. Hafner, K. Schindler, and V. Koltun, ``MiDaS: Towards High-Quality Monocular Depth Estimation,'' in Proc. IEEE Conf. Comput. Vis. Pattern Recognit. (CVPR) Workshops, 2020, pp. 131–144.
\bibitem{ref4} B. Robert, M. David, and A. Martin, ``Challenges in Tree Species Identification from Mid-Range Images: Dataset Interoperability and Standardization,'' Ecol. Inform., vol. 60, p. 101284, 2020.
\bibitem{ref5} N. Kumar, P. N. Belhumeur, A. Biswas, D. W. Jacobs, W. J. Kress, I. C. Lopez, and J. V. Soares, ``Leafsnap: A Computer Vision System for Automatic Plant Species Identification,'' in Proc. Eur. Conf. Comput. Vis. (ECCV), 2012, pp. 502–516.
\bibitem{ref6} H. Goëau, P. Bonnet, and A. Joly, ``Plant Identification in an Open World (Pl@ntNet),'' in Proc. Eur. Conf. Comput. Vis. Workshops (ECCVW), 2013, pp. 21–33.
\bibitem{ref7} [R. Cerutti, L. Tougne, A. Vacavant, and D. Coquin, ``Folia: An Application for Plant Identification Using Leaf Image Analysis,'' in Proc. Int. Conf. Image Anal. Recognit. (ICIAR), 2013, pp. 527–534.
\bibitem{ref8} H. Goëau, P. Bonnet, A. Joly, V. Bakić, J. Barbe, S. Selmi, I. Yahiaoui, and A. Barthélémy, ``Pl@ntLeaves: A Leaf Image Database,'' in Proc. Int. Conf. Multimedia Retr., 2012.
\bibitem{ref9} M. Carpentier, P. Giguère, and J. Gaudreault, ``BarkNet 1.0: A Dataset for Bark Classification,'' arXiv preprint arXiv:1806.03474, 2018.
\bibitem{ref10} M. Ratajczak, M. Kozioł, and A. Kozłowski, ``Bark-101: A Dataset for Bark Species Classification Using Deep Convolutional Neural Networks,'' in Proc. Int. Conf. Image Process. (ICIP), 2019, pp. 103–107.
\bibitem{ref11} T. Kim, Y. Zhang, and Q. Wang, ``UrbanStreetTree: A Large-Scale Dataset for Urban Tree Species Classification and Segmentation,'' Remote Sens., vol. 14, no. 4, p. 932, 2022.
\bibitem{ref12} K. He, X. Zhang, S. Ren, and J. Sun, ``Deep Residual Learning for Image Recognition,'' in Proc. IEEE Conf. Comput. Vis. Pattern Recognit. (CVPR), 2016, pp. 770–778.
\bibitem{ref13} M. Caron, H. Touvron, I. Misra, H. Jégou, J. Mairal, P. Bojanowski, and A. Joulin, ``Emerging Properties in Self-Supervised Vision Transformers,'' in Proc. IEEE Int. Conf. Comput. Vis. (ICCV), 2021, pp. 9650–9660.
\bibitem{ref14} Z. Liu et al., ``Swin Transformer: Hierarchical Vision Transformer using Shifted Windows,'' in Proc. IEEE Int. Conf. Comput. Vis. (ICCV), 2021, pp. 10012–10022.
\end{thebibliography}
\endgroup


\end{document}
